---
title: "Homework 6"
subtitle: "Autumn 2024"
author: "WRITE YOUR NAME HERE"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb, setspace}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---
\onehalfspacing

* * * 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = F, 
                      warning = F)

library(tidyverse)

```



### Instructions

- This homework is due in Gradescope on Wednesday Nov 20 by midnight PST. There is a 15 minute grace period and submissions made during this time will not be marked as late. Any work submitted past this period is considered late.  

- Please answer the following questions in the order in which they are posed. 

- Don't forget to (i) make a local copy this document for your work and to (ii) knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 

* * *



### Exercises 

1. (Difference of two means) Using a random sample of 25 observations - $X_1,X_2,\dots,X_{25}$ -  from a  population with (unknown) mean $\mu_1$ and (known) standard deviation $\sigma_1$, a 95\% large sample confidence interval for $\mu_1$ was found to be $[1.37, 1.53]$. 

   Similarly using an independent sample of 36 observations - $Y_1, Y_2, \dots, Y_{36}$ - from a population with (unknown) mean $\mu_2$ and (known) standard deviation $\sigma_2$, a 95\% large sample confidence interval for $\mu_2$ was found to be $[1.17, 1.29]$.

a. Find the values for $\bar{x}$, $\bar{y}$, $\sigma_1$ and $\sigma_2$. Be sure to 
   - write the formula for the 95\% interval estimate for a generic mean $\mu$ 
   - explain how you will find the values for $\bar{x}$ and $\sigma$ from your generic interval
   - then use R as a calculator to find and print the four values requested. 

b. The large sample confidence interval for a population mean is based on the Central Limit Theorem which states that sample means are approximately normally distributed for large means. Specifically, 
$$\bar{X} \approx Norm(\mu_1, \frac{\sigma_1}{\sqrt{25}}) \ \ \  \ \ \ \bar{Y} \approx Norm(\mu_2, \frac{\sigma_2}{\sqrt{36}}).$$
This implies that $\bar{X} - \bar{Y}$ is also approximately normally distributed for large sample sizes^[since linear combination of normal random variables is also normal]. What are the mean and standard deviation of this distribution? 


c. Use the large sample normal distribution of $\bar{X} - \bar{Y}$ you derived in part b derive a $100 ( 1-\alpha)$\% confidence interval for $\mu_1- \mu_2$. 

    Hint: You need to replace the standard normal random variable $Z$ in the probability statement
    $$1-\alpha = P\left( -z_{\alpha/2} \leq Z \leq z_{\alpha/2} \right)$$
    with the \emph{standardized} version of $\bar{X} - \bar{Y}$ and then invert the probability.
    
d. Using R as a calculator, find a 90\% confidence interval for this example data. Be sure to print the interval.


2. (Binomial probability) A sample of 83 independent realizations of an integer-valued random variable $X$ is shown below:
\begin{table}[h]
\centering
\begin{tabular}{ccccccc} \\ \hline
value & 0 & 1 & 2 & 3 & 4 & 5 \\ \hline
frequency & 13 & 18 & 23 & 15 & 6 & 8 \\ \hline
\end{tabular}
\end{table}

Suppose we are interested in estimating $\pi_0 = P(X \geq 2)$. Let the Bernoulli random variable $Y$ denote whether or not $X \geq 2$. Then $Y_1,Y_2,\dots,Y_{83} \overset{i.i.d.}{\sim} Bernoulli(\pi_0)$. 

    Calculate a 90% large sample confidence interval for $\pi_0$. You can simply state the formula for the confidence interval and then do the calculations in a code chunk using R as a calculator. Be sure to print your interval.
 

    
3. (Poisson rate) Suppose $X_1,X_2,\dots,X_n \overset{i.i.d.}{\sim} Pois(\lambda_0)$ where $\lambda_0$ is the true (but unknown ) value of the rate. 
a. What are the mean - $\mu_0$ - and variance  - $\sigma_0^2$ - for the Poisson distribution? (No proof needed)

b. What does the Central Limit Theorem say about the large sample distribution of $\bar{X}$ for the Poisson model? (No proof needed, just cite any references you use)

c. Write a definition like the one in "Definition 5.2" from $\S 5.4$ giving the formula for a large sample confidence interval for $\lambda_0$. (No proof needed)

    \begin{shaded}
    \textbf{Definition 5.2 (Poisson)}

    \end{shaded}

d. Suppose $X$, the number of students who come to class late can be modeled as a Poisson random variable with rate $\lambda_0$. The data are shown below for 12 classes in a frequency table. (For example, there were 4 classes with 0 late students). Use R to calculate a 98\% confidence interval estimate for $\lambda_0$. Be sure to print the answer.

    \begin{tabular}{cccccc} 
    number of late students & 0 & 1 & 2 & 3 & 4 \\ \hline
    frequency &              4 & 4 & 2 & 1 & 1  \\ \hline
    \end{tabular}
    
    Use R to calculate a 98\% confidence interval estimate for $\lambda_0$. Be sure to print the answer.
    ```{r 3d}
    #calculate 98% interval and print
    
    ```

4. (Bootstrap airbnb) In this problem we will revisit the `airbnb` data from section 5.3 to make inference about the **MEDIAN** (weekly) rental price of ALL homes in Seattle with 3 or more bedrooms. Since we are interested in estimating the MEDIAN - let's call it $\mu_d$ - of a distribution, we will use the sample median as our estimate.  In this exercise, you will calculate a bootstrapped interval estimate for $\mu_d$.

a. Read in the data from the file `listings.csv` into a dataframe called `airbnb`. Then write code to create a smaller data frame called `airbnb_3bed` which contains only the variable `weekly_price` (as a number) for all Houses with 3 or more bedrooms. End the code chunk with a `glimpse` so we can peek at the data. We do not need to see your code which is why I have set `echo = FALSE`.

    Hint: The variables you will be filtering on are given below. 

    - property_type: House
    - room_type: Entire home/apt
    - bedrooms: 3 or more 
  
    ```{r problem 4a, echo = F}
    #create airbnb_3bed

    # glimpse
    ```


b. Clearly the weekly price is missing or NA for many listings. Write code to filter out the missing values and save the result in `airbnb_3bed`. Print the number of rows in the resulting dataframe. Also calculate and print the median of `weekly_price`. This is our observed sample estimate for $\mu_d$


    ```{r problem 4b}
    #filter to remove missing values of weekly_price 

    

    ```


c. In this part, you will use your dataframe `airbnb_3bed` from part b. to construct a 95% (non-parametric) bootstrap percentile interval estimate for $\mu_{50}$ - the median price of all House rentals in Seattle in 2018 with 3 or more bedrooms. 

    We are only grading this for accuracy of the numbers reported in the table below.  

   
    ```{r problem4c}

    #Generate sample of n = nrow(airbnb_3bed) from airbnb_3bed dataframe
    #with replacement
    #Calculate m_50_hat = median(weekly_price) from each sample.
    #Repeat B times
    #Calculate average and sd of the B values 


    set.seed(1141)
    B = 1000
    ```
  
  
  
\begin{table}[h]
\centering
\caption{Bootstrapping Results (rounded to 3 decimal places) for estimating $\mu_{50}$ }
\begin{tabular}{|p{4in}|p{1in}| } \hline
Average of bootstrapped estimates of $\mu_{50}$ &  \\
S.D. of bootstrapped estimates of $\mu_{50}$ & \\
Bootstrap estimate of bias in the median as an estimator & \\
Bias-corrected sample estimate of $\mu_{50}$ & \\
Lower end point of 95\% bootstrap interval for $\mu_{50}$ & \\
Upper end point of 95\% bootstrap interval for $\mu_{50}$ & \\ \hline
\end{tabular}
\end{table}
